---
title: "Apple Watch Data Conversion"
output: 
  html_document:
    theme: yeti
    highlight: kate
    code_folding: hide
---

```{r, warning=FALSE, message=FALSE}
#load libaries
library(pander)
library(magrittr)
library(dplyr)
library(xml2)
library(tidyverse)
library(readr)
library(XML)
library(car)
library(DT)
library(shiny)
library(lubridate)
```

# Data Size and Import

There is a lot of data in this file, so we first want to see how many records there are in this set. 

```{r}
#import apple watch data file
##This should be an xml file, if the xml file won't run properly then ensure that your iphone is up to date and redownload the health data.
records <- read_xml("exportholly.xml") %>%
  xml_children

#View record length
pander(length(records))
```

From this we want to see how many records we have of each type of data collected off of the apple watch. Since some of the types of records have such a low amount, they may not be helpful.

```{r}
#Export record types
record_types <- xml_attr(records, "type")

#view table of record types that are pulled from the xml file
record_types %>%
  table(dnn="Record_type") %>%            
  as.data.frame %>%
  arrange(desc(Freq)) %>%
  mutate(Count = prettyNum(Freq, big.mark=",")) %>%
  select(-Freq) %>%
  pander()
```

# Pulling out Data Types

For the xml file, I have to pull out all of the records separately and join them together later. You will have to pick from the above types and follow the code below to pull out that data. You may need to change some of the code if you are not pulling out the same records.


```{r, warning=FALSE, message=FALSE}

#######

#Extract Heart Rate Records
heart_rate_records <- records[which(record_types == "HKQuantityTypeIdentifierHeartRate")]

#Heart Rate Dataframe
heart_rate_df <- data_frame(date = strptime(xml_attr(heart_rate_records, "endDate"), '%Y-%m-%d %H:%M:%S %z'),
                            heart_rate = as.integer(xml_attr(heart_rate_records, "value")))

#Fix the date so that the format can join our records later
heart_rate_df$date <- ymd_hms(heart_rate_df$date)

#all of the data is at a minute in time, we would have a lot of NAs if we kept to minutes so this changes it to take the average for the hour, this could also be changed to day if that would be better for your data.
Heart_Rate <- heart_rate_df %>%
  group_by(date=floor_date(date,'hour')) %>%
  summarize(Mean_Heart_Rate=mean(heart_rate))


#######

#Extract active energy Records
active_energy_records <- records[which(record_types == "HKQuantityTypeIdentifierActiveEnergyBurned")]

#Active Energy Data frame
active_energy_df <- data_frame(date = strptime(xml_attr(active_energy_records, "endDate"), '%Y-%m-%d %H:%M:%S %z'),
                            active_energy = as.integer(xml_attr(active_energy_records, "value")))

active_energy_df$date <- ymd_hms(active_energy_df$date)

Active_Energy <- active_energy_df %>%
  group_by(date=floor_date(date,'hour')) %>%
  summarize(Mean_Active_Energy=mean(active_energy))


#######

#Extract basal energy Records
basal_energy_records <- records[which(record_types == "HKQuantityTypeIdentifierBasalEnergyBurned")]

#Basal Energy Data frame
basal_energy_df <- data_frame(date = strptime(xml_attr(basal_energy_records, "endDate"), '%Y-%m-%d %H:%M:%S %z'),
                            basal_energy = as.integer(xml_attr(basal_energy_records, "value")))

basal_energy_df$date <- ymd_hms(basal_energy_df$date)

Basal_Energy <- basal_energy_df %>%
  group_by(date=floor_date(date,'hour')) %>%
  summarize(Mean_Basal_Energy=mean(basal_energy))


#######

#Extract Distance Records
distance_records <- records[which(record_types == "HKQuantityTypeIdentifierDistanceWalkingRunning")]


#Distance Data frame
distance_df <- data_frame(date = strptime(xml_attr(distance_records, "endDate"), '%Y-%m-%d %H:%M:%S %z'),
                            distance = as.integer(xml_attr(distance_records, "value")))

distance_df$date <- ymd_hms(distance_df$date)

Distance <- distance_df %>%
  group_by(date=floor_date(date,'hour')) %>%
  summarize(Distance=sum(distance))


#######

#Extract steps records
steps_records <- records[which(record_types == "HKQuantityTypeIdentifierStepCount")]

#Steps Data frame
steps_df <- data_frame(date = strptime(xml_attr(steps_records, "endDate"), '%Y-%m-%d %H:%M:%S %z'),
                            steps = as.integer(xml_attr(steps_records, "value")))

steps_df$date <- ymd_hms(steps_df$date)

Steps <- steps_df %>%
  group_by(date=floor_date(date,'hour')) %>%
  summarize(Steps=sum(steps))


#######

#Extract Stand Time Records
stand_time_records <- records[which(record_types == "HKQuantityTypeIdentifierAppleStandTime")]

#Stand Time Data frame
stand_time_df <- data_frame(date = strptime(xml_attr(stand_time_records, "endDate"), '%Y-%m-%d %H:%M:%S %z'),
                            stand_time = as.integer(xml_attr(stand_time_records, "value")))

stand_time_df$date <- ymd_hms(stand_time_df$date)

Stand_Time <- stand_time_df %>%
  group_by(date=floor_date(date,'hour')) %>%
  summarize(Stand_Time=sum(stand_time))


#######

#Extract walking speed Records
walking_speed_records <- records[which(record_types == "HKQuantityTypeIdentifierWalkingSpeed")]

#Walking Speed Data frame
walking_speed_df <- data_frame(date = strptime(xml_attr(walking_speed_records, "endDate"), '%Y-%m-%d %H:%M:%S %z'),
                            walking_speed = as.integer(xml_attr(walking_speed_records, "value")))

walking_speed_df$date <- ymd_hms(walking_speed_df$date)

Walking_Speed <- walking_speed_df %>%
  group_by(date=floor_date(date,'hour')) %>%
  summarize(Mean_Walking_Speed=mean(walking_speed))


#######

#Extract Exercise Time Records
exercise_time_records <- records[which(record_types == "HKQuantityTypeIdentifierAppleExerciseTime")]

#Exercise Time Data frame
exercise_time_df <- data_frame(date = strptime(xml_attr(exercise_time_records, "endDate"), '%Y-%m-%d %H:%M:%S %z'),
                            exercise_time = as.integer(xml_attr(exercise_time_records, "value")))

exercise_time_df$date <- ymd_hms(exercise_time_df$date)

Exercise_Time <- exercise_time_df %>%
  group_by(date=floor_date(date,'hour')) %>%
  summarize(Exercise_Time=sum(exercise_time))

#######

#Extract Stair Flights Records
stairs_records <- records[which(record_types == "HKQuantityTypeIdentifierFlightsClimbed")]

#Stair Data frame
stairs_df <- data_frame(date = strptime(xml_attr(stairs_records, "endDate"), '%Y-%m-%d %H:%M:%S %z'),
                            stairs = as.integer(xml_attr(stairs_records, "value")))

stairs_df$date <- ymd_hms(stairs_df$date)

Stairs <- stairs_df %>%
  group_by(date=floor_date(date,'hour')) %>%
  summarize(Stairs=sum(stairs))

#######

#Extract Step Length Records
step_length_records <- records[which(record_types == "HKQuantityTypeIdentifierWalkingStepLength")]

#Step Length Data frame
step_length_df <- data_frame(date = strptime(xml_attr(step_length_records, "endDate"), '%Y-%m-%d %H:%M:%S %z'),
                            step_length = as.integer(xml_attr(step_length_records, "value")))

step_length_df$date <- ymd_hms(step_length_df$date)

Step_Length <- step_length_df %>%
  group_by(date=floor_date(date,'hour')) %>%
  summarize(Step_Length=mean(step_length))

#######

#Extract Walking Double Support Records
walk_double_records <- records[which(record_types == "HKQuantityTypeIdentifierWalkingDoubleSupportPercentage")]

#Walking Double Support Data frame
walk_double_df <- data_frame(date = strptime(xml_attr(walk_double_records, "endDate"), '%Y-%m-%d %H:%M:%S %z'),
                            walk_double = as.integer(xml_attr(walk_double_records, "value")))

walk_double_df$date <- ymd_hms(walk_double_df$date)

Walk_Double <- walk_double_df %>%
  group_by(date=floor_date(date,'hour')) %>%
  summarize(Walk_Double=mean(walk_double))

#######

#Extract Walking Asymmetry Records
walk_asymmetry_records <- records[which(record_types == "HKQuantityTypeIdentifierWalkingAsymmetryPercentage")]

#Walking Asymmetry Data frame
walk_asymmetry_df <- data_frame(date = strptime(xml_attr(walk_asymmetry_records, "endDate"), '%Y-%m-%d %H:%M:%S %z'),
                            walk_asymmetry = as.integer(xml_attr(walk_asymmetry_records, "value")))

walk_asymmetry_df$date <- ymd_hms(walk_asymmetry_df$date)

Walk_Asymmetry <- walk_asymmetry_df %>%
  group_by(date=floor_date(date,'hour')) %>%
  summarize(Walk_Asymmetry=mean(walk_asymmetry))

#######

#Extract Sleep Analysis Support Records
sleep_records <- records[which(record_types == "HKCategoryTypeIdentifierSleepAnalysis")]

#Sleep Data frame
sleep_df <- data_frame(date = strptime(xml_attr(sleep_records, "endDate"), '%Y-%m-%d %H:%M:%S %z'),
                            sleep = as.integer(xml_attr(sleep_records, "value")))

sleep_df$date <- ymd_hms(sleep_df$date)

Sleep <- sleep_df %>%
  group_by(date=floor_date(date,'hour')) %>%
  summarize(Sleep=mean(sleep))

#######

#Extract Heart Rate Variability Records
heart_rate_variability_records <- records[which(record_types == "HKQuantityTypeIdentifierHeartRateVariabilitySDNN")]

#Heart Rate Variability Data frame
heart_rate_variability_df <- data_frame(date = strptime(xml_attr(heart_rate_variability_records, "endDate"), '%Y-%m-%d %H:%M:%S %z'),
                            heart_rate_variability = as.integer(xml_attr(heart_rate_variability_records, "value")))

heart_rate_variability_df$date <- ymd_hms(heart_rate_variability_df$date)

Heart_Rate_Variability <- heart_rate_variability_df %>%
  group_by(date=floor_date(date,'hour')) %>%
  summarize(Heart_Rate_Variability=mean(heart_rate_variability))

#######

#Extract Environment Audio Exposure Records
enviornment_audio_exposure_records <- records[which(record_types == "HKQuantityTypeIdentifierEnvironmentalAudioExposure")]

#Environment Audio Exposure Data frame
enviornment_audio_exposure_df <- data_frame(date = strptime(xml_attr(enviornment_audio_exposure_records, "endDate"), '%Y-%m-%d %H:%M:%S %z'),
                            enviornment_audio_exposure = as.integer(xml_attr(enviornment_audio_exposure_records, "value")))

enviornment_audio_exposure_df$date <- ymd_hms(enviornment_audio_exposure_df$date)

Enviornment_Audio_Exposure <- enviornment_audio_exposure_df %>%
  group_by(date=floor_date(date,'hour')) %>%
  summarize(Enviornment_Audio_Exposure=mean(enviornment_audio_exposure))

#######

#Extract Headphone Audio Exposure Records
headphone_audio_exposure_records <- records[which(record_types == "HKQuantityTypeIdentifierHeadphoneAudioExposure")]

#Headphone Audio Exposure Data frame
headphone_audio_exposure_df <- data_frame(date = strptime(xml_attr(headphone_audio_exposure_records, "endDate"), '%Y-%m-%d %H:%M:%S %z'),
                            headphone_audio_exposure = as.integer(xml_attr(headphone_audio_exposure_records, "value")))

headphone_audio_exposure_df$date <- ymd_hms(headphone_audio_exposure_df$date)

Headphone_Audio_Exposure <- headphone_audio_exposure_df %>%
  group_by(date=floor_date(date,'hour')) %>%
  summarize(Headphone_Audio_Exposure=mean(headphone_audio_exposure))

#######

#Extract Stair Ascent Speed Records
stair_ascent_speed_records <- records[which(record_types == "HKQuantityTypeIdentifierStairAscentSpeed")]

#Stair Ascent Speed Data frame
stair_ascent_speed_df <- data_frame(date = strptime(xml_attr(stair_ascent_speed_records, "endDate"), '%Y-%m-%d %H:%M:%S %z'),
                            stair_ascent_speed = as.integer(xml_attr(stair_ascent_speed_records, "value")))

stair_ascent_speed_df$date <- ymd_hms(stair_ascent_speed_df$date)

Stair_Ascent_Speed <- stair_ascent_speed_df %>%
  group_by(date=floor_date(date,'hour')) %>%
  summarize(Stair_Ascent_Speed=mean(stair_ascent_speed))

#######

#Extract Stair Descent Speed Records
stair_descent_speed_records <- records[which(record_types == "HKQuantityTypeIdentifierStairDescentSpeed")]

#Stair Descent Speed Data frame
stair_descent_speed_df <- data_frame(date = strptime(xml_attr(stair_descent_speed_records, "endDate"), '%Y-%m-%d %H:%M:%S %z'),
                            stair_descent_speed = as.integer(xml_attr(stair_descent_speed_records, "value")))

stair_descent_speed_df$date <- ymd_hms(stair_descent_speed_df$date)

Stair_Descent_Speed <- stair_descent_speed_df %>%
  group_by(date=floor_date(date,'hour')) %>%
  summarize(Stair_Descent_Speed=mean(stair_descent_speed))

#######

#Extract Resting Heart Rate Records
resting_heart_rate_records <- records[which(record_types == "HKQuantityTypeIdentifierRestingHeartRate")]

#Resting Heart Rate Data frame
resting_heart_rate_df <- data_frame(date = strptime(xml_attr(resting_heart_rate_records, "endDate"), '%Y-%m-%d %H:%M:%S %z'),
                            resting_heart_rate = as.integer(xml_attr(resting_heart_rate_records, "value")))

resting_heart_rate_df$date <- ymd_hms(resting_heart_rate_df$date)

Resting_Heart_Rate <- resting_heart_rate_df %>%
  group_by(date=floor_date(date,'hour')) %>%
  summarize(Resting_Heart_Rate=mean(resting_heart_rate))

#######

#Extract Walking Heart Rate Records
walking_heart_rate_records <- records[which(record_types == "HKQuantityTypeIdentifierWalkingHeartRateAverage")]

#Walking Heart Rate Data frame
walking_heart_rate_df <- data_frame(date = strptime(xml_attr(walking_heart_rate_records, "endDate"), '%Y-%m-%d %H:%M:%S %z'),
                            walking_heart_rate = as.integer(xml_attr(walking_heart_rate_records, "value")))

walking_heart_rate_df$date <- ymd_hms(walking_heart_rate_df$date)

Walking_Heart_Rate <- walking_heart_rate_df %>%
  group_by(date=floor_date(date,'hour')) %>%
  summarize(Walking_Heart_Rate=mean(walking_heart_rate))

#######

#Extract Respiratory Rate Records
respiratory_rate_records <- records[which(record_types == "HKQuantityTypeIdentifierRespiratoryRate")]

#Respiratory Rate Data frame
respiratory_rate_df <- data_frame(date = strptime(xml_attr(respiratory_rate_records, "endDate"), '%Y-%m-%d %H:%M:%S %z'),
                            respiratory_rate = as.integer(xml_attr(respiratory_rate_records, "value")))

respiratory_rate_df$date <- ymd_hms(respiratory_rate_df$date)

Respiratory_Rate <- respiratory_rate_df %>%
  group_by(date=floor_date(date,'hour')) %>%
  summarize(Respiratory_Rate=mean(respiratory_rate))

#######

#Extract Hand Wash Records
hand_wash_records <- records[which(record_types == "HKCategoryTypeIdentifierHandwashingEvent")]

#Hand Wash Data frame
hand_wash_df <- data_frame(date = strptime(xml_attr(hand_wash_records, "endDate"), '%Y-%m-%d %H:%M:%S %z'),
                            hand_wash = as.integer(xml_attr(hand_wash_records, "value")))

hand_wash_df$date <- ymd_hms(hand_wash_df$date)

Hand_Wash <- hand_wash_df %>%
  group_by(date=floor_date(date,'hour')) %>%
  summarize(Hand_Wash=mean(hand_wash))

#######

#Extract Walk Steadiness Records
walk_steadiness_records <- records[which(record_types == "HKQuantityTypeIdentifierAppleWalkingSteadiness")]

#Walk Steadiness Data frame
walk_steadiness_df <- data_frame(date = strptime(xml_attr(walk_steadiness_records, "endDate"), '%Y-%m-%d %H:%M:%S %z'),
                            walk_steadiness = as.integer(xml_attr(walk_steadiness_records, "value")))

walk_steadiness_df$date <- ymd_hms(walk_steadiness_df$date)

Walk_Steadiness <- walk_steadiness_df %>%
  group_by(date=floor_date(date,'hour')) %>%
  summarize(Walk_Steadiness=mean(walk_steadiness))

#######

#Extract Body Mass Records
body_mass_records <- records[which(record_types == "HKQuantityTypeIdentifierBodyMass")]

#Body Mass Data frame
body_mass_df <- data_frame(date = strptime(xml_attr(body_mass_records, "endDate"), '%Y-%m-%d %H:%M:%S %z'),
                            body_mass = as.integer(xml_attr(body_mass_records, "value")))

body_mass_df$date <- ymd_hms(body_mass_df$date)

Body_Mass <- body_mass_df %>%
  group_by(date=floor_date(date,'hour')) %>%
  summarize(Body_Mass=mean(body_mass))

#######

#Extract VO2 Max Records
vo2_max_records <- records[which(record_types == "HKQuantityTypeIdentifierVO2Max")]

#Vo2 Max Data frame
vo2_max_df <- data_frame(date = strptime(xml_attr(vo2_max_records, "endDate"), '%Y-%m-%d %H:%M:%S %z'),
                            vo2_max = as.integer(xml_attr(vo2_max_records, "value")))

vo2_max_df$date <- ymd_hms(vo2_max_df$date)

Vo2_Max <- vo2_max_df %>%
  group_by(date=floor_date(date,'hour')) %>%
  summarize(Vo2_Max=mean(vo2_max))

#######

#Extract Swim Distance Records
swim_distance_records <- records[which(record_types == "HKQuantityTypeIdentifierDistanceSwimming")]

#Hand Wash Data frame
swim_distance_df <- data_frame(date = strptime(xml_attr(swim_distance_records, "endDate"), '%Y-%m-%d %H:%M:%S %z'),
                            swim_distance = as.integer(xml_attr(swim_distance_records, "value")))

swim_distance_df$date <- ymd_hms(swim_distance_df$date)

Swim_Distance <- swim_distance_df %>%
  group_by(date=floor_date(date,'hour')) %>%
  summarize(Swim_Distance=sum(swim_distance))

#######

#Extract Swimming Stroke Count Records
swim_stroke_records <- records[which(record_types == "HKQuantityTypeIdentifierSwimmingStrokeCount")]

#Swim Stroke Data frame
swim_stroke_df <- data_frame(date = strptime(xml_attr(swim_stroke_records, "endDate"), '%Y-%m-%d %H:%M:%S %z'),
                            swim_stroke = as.integer(xml_attr(swim_stroke_records, "value")))

swim_stroke_df$date <- ymd_hms(swim_stroke_df$date)

Swim_Stroke <- swim_stroke_df %>%
  group_by(date=floor_date(date,'hour')) %>%
  summarize(Swim_Stroke=sum(swim_stroke))

#######

#Extract Mindful Records
mindful_records <- records[which(record_types == "HKCategoryTypeIdentifierMindfulSession")]

#Mindful Data frame
mindful_df <- data_frame(date = strptime(xml_attr(mindful_records, "endDate"), '%Y-%m-%d %H:%M:%S %z'),
                            mindful = as.integer(xml_attr(mindful_records, "value")))

mindful_df$date <- ymd_hms(mindful_df$date)

Mindful <- mindful_df %>%
  group_by(date=floor_date(date,'hour')) %>%
  summarize(Mindful=mean(mindful))

#######

#Extract Dietary Water Records
dietary_water_records <- records[which(record_types == "HKQuantityTypeIdentifierDietaryWater")]

#Dietary Water Data frame
dietary_water_df <- data_frame(date = strptime(xml_attr(dietary_water_records, "endDate"), '%Y-%m-%d %H:%M:%S %z'),
                            dietary_water = as.integer(xml_attr(dietary_water_records, "value")))

dietary_water_df$date <- ymd_hms(dietary_water_df$date)

Dietary_Water <- dietary_water_df %>%
  group_by(date=floor_date(date,'hour')) %>%
  summarize(Dietary_Water=mean(dietary_water))

#######

#Extract Cycling Records
cycling_records <- records[which(record_types == "HKQuantityTypeIdentifierDistanceCycling")]

#Cycling Data frame
cycling_df <- data_frame(date = strptime(xml_attr(cycling_records, "endDate"), '%Y-%m-%d %H:%M:%S %z'),
                            cycling = as.integer(xml_attr(cycling_records, "value")))

cycling_df$date <- ymd_hms(cycling_df$date)

Cycling <- cycling_df %>%
  group_by(date=floor_date(date,'hour')) %>%
  summarize(Cycling=sum(cycling))
```


## Combine all Data Sets

This next part combines all of the data sets and saves it to a csv file that can be used for 

```{r, warning=FALSE, message=FALSE}
#Join all data
AppleWatch <- left_join(Heart_Rate, Active_Energy, by='date')
AppleWatch <- left_join(AppleWatch, Basal_Energy, by='date')
AppleWatch <- left_join(AppleWatch, Distance, by='date')
AppleWatch <- left_join(AppleWatch, Steps, by='date')
AppleWatch <- left_join(AppleWatch, Stand_Time, by='date')
AppleWatch <- left_join(AppleWatch, Stairs, by='date')
AppleWatch <- left_join(AppleWatch, Exercise_Time, by='date')
AppleWatch <- left_join(AppleWatch, Walking_Speed, by='date')
AppleWatch <- left_join(AppleWatch, Walk_Double, by='date')
AppleWatch <- left_join(AppleWatch, Walk_Asymmetry, by='date')
AppleWatch <- left_join(AppleWatch, Sleep, by='date')
AppleWatch <- left_join(AppleWatch, Heart_Rate_Variability, by='date')
AppleWatch <- left_join(AppleWatch, Enviornment_Audio_Exposure, by='date')
AppleWatch <- left_join(AppleWatch, Headphone_Audio_Exposure, by='date')
AppleWatch <- left_join(AppleWatch, Stair_Ascent_Speed, by='date')
AppleWatch <- left_join(AppleWatch, Stair_Descent_Speed, by='date')
AppleWatch <- left_join(AppleWatch, Resting_Heart_Rate, by='date')
AppleWatch <- left_join(AppleWatch, Walking_Heart_Rate, by='date')
AppleWatch <- left_join(AppleWatch, Respiratory_Rate, by='date')
AppleWatch <- left_join(AppleWatch, Hand_Wash, by='date')
AppleWatch <- left_join(AppleWatch, Walk_Steadiness, by='date')
AppleWatch <- left_join(AppleWatch, Body_Mass, by='date')
AppleWatch <- left_join(AppleWatch, Vo2_Max, by='date')
AppleWatch <- left_join(AppleWatch, Swim_Distance, by='date')
AppleWatch <- left_join(AppleWatch, Swim_Stroke, by='date')
AppleWatch <- left_join(AppleWatch, Mindful, by='date')
AppleWatch <- left_join(AppleWatch, Dietary_Water, by='date')
AppleWatch <- left_join(AppleWatch, Cycling, by='date')

#This rounds the decimal places since there are a lot of digits
AppleWatch <- AppleWatch %>%
  mutate_if(is.numeric, round, digits=2)

#view complete dataset
datatable(AppleWatch)

#save as csv file
##You can access the file through whatever path you input down below
write.csv(AppleWatch, '/Users/jillh/Documents/Senior_Project/AppleWatch.csv', row.names=FALSE)

```