---
title: "Data Exploration and Cleanup"
output: 
  html_document:
    theme: yeti
    highlight: kate
    code_folding: hide
---

```{r, warning=FALSE, message=FALSE}
#load libaries
library(pander)
library(magrittr)
library(dplyr)
library(xml2)
library(tidyverse)
library(readr)
library(XML)
library(car)
library(DT)
library(shiny)
library(lubridate)
```

## Dexcom Data

The Dexcom G6 Continuous Glucose Monitoring (CGM) System sends Glucose numbers to a smart phone every 5 minutes. 

This data was collected for May 1st, 2022 - April 30th, 2023. The following code sorts this data and cleans it up, the following data table are the resulting glucose numbers.

```{r, warning=FALSE, message=FALSE}
#Dexcom Data Loading
Dexcom1 <- read_csv("Dexcom1.csv")
Dexcom2 <- read_csv("Dexcom2.csv")
Dexcom3 <- read_csv("Dexcom3.csv")
Dexcom4 <- read_csv("Dexcom4.csv")
Dexcom5 <- read_csv("Dexcom5.csv")

#Join Dexcom Data Files
Dexcom <- rbind(Dexcom1, Dexcom2, Dexcom3, Dexcom4, Dexcom5)

#Mutating Column Names
Dex_Clean <- Dexcom %>%
  mutate(Date =`Timestamp (YYYY-MM-DDThh:mm:ss)`,
         Event=`Event Type`,
         Sub_Event = `Event Subtype`,
         Glucose = `Glucose Value (mg/dL)`)

#Replace 'High' in glucose column with value of 400 (Information came from a key where data was collected)
Dex_Clean$Glucose <- replace(Dex_Clean$Glucose, Dex_Clean$Glucose=="High", 400)

#Take only the needed columns
Dex_Clean <- select(Dex_Clean, c(Date, Glucose))

#Reformat the date column so it can be joined later
Dex_Clean$Date <- ymd_hms(Dex_Clean$Date)
Dex_Clean$Glucose <- as.numeric(Dex_Clean$Glucose)

#Take mean glucose level for each hour so it will match other data set formats
Dex_Final <- Dex_Clean %>%
  group_by(date=floor_date(Date,'hour')) %>%
  summarize(Mean_Glucose=mean(Glucose))

#View Dex data
datatable(Dex_Clean)
```

## Weather Data

I imported weather data to also see how blood sugar levels compare to the weather and to test if that would have any impact, the following table shows the weather data that was collected for the same time frame as the Dexcom Data.

```{r, warning=FALSE, message=FALSE}
#Import weather data
Weather <- read_csv("dataexport_20230509T201314.csv")

#View weather data
datatable(Weather)
```

## Health Data

Health data was taken from an apple watch to also compare how that influences with blood sugar levels, we will look at all of these separately. 

#### Load Data

There is a lot of data in this file, so we first want to see how many records there are in this set. 

```{r}
#import apple watch data file
records <- read_xml("exportholly.xml") %>%
  xml_children

#View record length
pander(length(records))
```

From this we want to see how many records we have of each type of data collected off of the apple watch. Since some of the types of records have such a low amount, and we are trying to compare this to blood sugar data, not all of these types will be used.

```{r}
#Export record types
record_types <- xml_attr(records, "type")

#view table of record types that are pulled from the xml file
record_types %>%
  table(dnn="Record_type") %>%            
  as.data.frame %>%
  arrange(desc(Freq)) %>%
  mutate(Count = prettyNum(Freq, big.mark=",")) %>%
  select(-Freq) %>%
  pander()
```

#### Getting Data for Health Data

For the xml file, I have to pull out all of the records separately and join them together later. I picked the types from up above that I thought would have the most impact and had enough data to be significant. I also saved each record type into a dataframe that could be used later if wanted.

```{r, warning=FALSE, message=FALSE}

#######

#Extract Heart Rate Recrods
heart_rate_records <- records[which(record_types == "HKQuantityTypeIdentifierHeartRate")]

single_record_attrs <- xml_attrs(heart_rate_records[10000])[[1]]
##class(single_record_attrs)
##data_frame(attribute = names(single_record_attrs),value= single_record_attrs)


#Heart Rate Dataframe
heart_rate_df <- data_frame(date = strptime(xml_attr(heart_rate_records, "endDate"), '%Y-%m-%d %H:%M:%S %z'),
                            heart_rate = as.integer(xml_attr(heart_rate_records, "value")))

heart_rate_df$date <- ymd_hms(heart_rate_df$date)

Heart_Rate <- heart_rate_df %>%
  group_by(date=floor_date(date,'hour')) %>%
  summarize(Mean_Heart_Rate=mean(heart_rate))

saveRDS(Heart_Rate, file="heart_rate_df.rds")

#######

#Extract active energy Records
active_energy_records <- records[which(record_types == "HKQuantityTypeIdentifierActiveEnergyBurned")]

single_record_attrs <- xml_attrs(active_energy_records[10000])[[1]]
##class(single_record_attrs)
##data_frame(attribute = names(single_record_attrs), value=single_record_attrs)

#Active Energy Data frame
active_energy_df <- data_frame(date = strptime(xml_attr(active_energy_records, "endDate"), '%Y-%m-%d %H:%M:%S %z'),
                            active_energy = as.integer(xml_attr(active_energy_records, "value")))

active_energy_df$date <- ymd_hms(active_energy_df$date)

Active_Energy <- active_energy_df %>%
  group_by(date=floor_date(date,'hour')) %>%
  summarize(Mean_Active_Energy=mean(active_energy))

saveRDS(Active_Energy, file="active_energy_df.rds")

#######

#Extract basal energy Records
basal_energy_records <- records[which(record_types == "HKQuantityTypeIdentifierBasalEnergyBurned")]

single_record_attrs <- xml_attrs(basal_energy_records[10000])[[1]]
##class(single_record_attrs)
##data_frame(attribute = names(single_record_attrs),value= single_record_attrs)

#Basal Energy Data frame
basal_energy_df <- data_frame(date = strptime(xml_attr(basal_energy_records, "endDate"), '%Y-%m-%d %H:%M:%S %z'),
                            basal_energy = as.integer(xml_attr(basal_energy_records, "value")))

basal_energy_df$date <- ymd_hms(basal_energy_df$date)

Basal_Energy <- basal_energy_df %>%
  group_by(date=floor_date(date,'hour')) %>%
  summarize(Mean_Basal_Energy=mean(basal_energy))

saveRDS(Basal_Energy, file="Basal_Energy.rds")

#######

#Extract Distance Records
distance_records <- records[which(record_types == "HKQuantityTypeIdentifierDistanceWalkingRunning")]

single_record_attrs <- xml_attrs(distance_records[10000])[[1]]
##class(single_record_attrs)
##data_frame(attribute = names(single_record_attrs),value= single_record_attrs)

#Distance Data frame
distance_df <- data_frame(date = strptime(xml_attr(distance_records, "endDate"), '%Y-%m-%d %H:%M:%S %z'),
                            distance = as.integer(xml_attr(distance_records, "value")))

distance_df$date <- ymd_hms(distance_df$date)

Distance <- distance_df %>%
  group_by(date=floor_date(date,'hour')) %>%
  summarize(Distance=sum(distance))

saveRDS(Distance, file="distance_df.rds")

#######

#Extract steps records
steps_records <- records[which(record_types == "HKQuantityTypeIdentifierStepCount")]

single_record_attrs <- xml_attrs(steps_records[10000])[[1]]
##class(single_record_attrs)
##data_frame(attribute = names(single_record_attrs),value = single_record_attrs)

#Steps Data frame
steps_df <- data_frame(date = strptime(xml_attr(steps_records, "endDate"), '%Y-%m-%d %H:%M:%S %z'),
                            steps = as.integer(xml_attr(steps_records, "value")))

steps_df$date <- ymd_hms(steps_df$date)

Steps <- steps_df %>%
  group_by(date=floor_date(date,'hour')) %>%
  summarize(Steps=sum(steps))

saveRDS(Steps, file="steps_df.rds")

#######

#Extract Stand Time Records
stand_time_records <- records[which(record_types == "HKQuantityTypeIdentifierAppleStandTime")]

single_record_attrs <- xml_attrs(stand_time_records[10000])[[1]]
##class(single_record_attrs)
##data_frame(attribute = names(single_record_attrs),value= single_record_attrs)

#Stand Time Data frame
stand_time_df <- data_frame(date = strptime(xml_attr(stand_time_records, "endDate"), '%Y-%m-%d %H:%M:%S %z'),
                            stand_time = as.integer(xml_attr(stand_time_records, "value")))

stand_time_df$date <- ymd_hms(stand_time_df$date)

Stand_Time <- stand_time_df %>%
  group_by(date=floor_date(date,'hour')) %>%
  summarize(Stand_Time=sum(stand_time))

saveRDS(Stand_Time, file="stand_time_df.rds")

#######

#Extract walking speed Records
walking_speed_records <- records[which(record_types == "HKQuantityTypeIdentifierWalkingSpeed")]

single_record_attrs <- xml_attrs(walking_speed_records[10000])[[1]]
##class(single_record_attrs)

##data_frame(attribute = names(single_record_attrs),value = single_record_attrs)

#Walking Speed Data frame
walking_speed_df <- data_frame(date = strptime(xml_attr(walking_speed_records, "endDate"), '%Y-%m-%d %H:%M:%S %z'),
                            walking_speed = as.integer(xml_attr(walking_speed_records, "value")))

walking_speed_df$date <- ymd_hms(walking_speed_df$date)

Walking_Speed <- walking_speed_df %>%
  group_by(date=floor_date(date,'hour')) %>%
  summarize(Mean_Walking_Speed=mean(walking_speed))

saveRDS(Walking_Speed, file="walking_speed_df.rds")

#######

#Extract Exercise Time Records
exercise_time_records <- records[which(record_types == "HKQuantityTypeIdentifierAppleExerciseTime")]

single_record_attrs <- xml_attrs(exercise_time_records[800])[[1]]
##class(single_record_attrs)

##data_frame(attribute = names(single_record_attrs),value= single_record_attrs)

#Exercise Time Data frame
exercise_time_df <- data_frame(date = strptime(xml_attr(exercise_time_records, "endDate"), '%Y-%m-%d %H:%M:%S %z'),
                            exercise_time = as.integer(xml_attr(exercise_time_records, "value")))

exercise_time_df$date <- ymd_hms(exercise_time_df$date)

Exercise_Time <- exercise_time_df %>%
  group_by(date=floor_date(date,'hour')) %>%
  summarize(Exercise_Time=sum(exercise_time))

saveRDS(Exercise_Time, file="exercise_time_df.rds")

#######

#Extract Stair Flights Records
stairs_records <- records[which(record_types == "HKQuantityTypeIdentifierFlightsClimbed")]

single_record_attrs <- xml_attrs(stairs_records[800])[[1]]
##class(single_record_attrs)

##data_frame(attribute = names(single_record_attrs),value= single_record_attrs)

#Stair Data frame
stairs_df <- data_frame(date = strptime(xml_attr(stairs_records, "endDate"), '%Y-%m-%d %H:%M:%S %z'),
                            stairs = as.integer(xml_attr(stairs_records, "value")))

stairs_df$date <- ymd_hms(stairs_df$date)

Stairs <- stairs_df %>%
  group_by(date=floor_date(date,'hour')) %>%
  summarize(Stairs=mean(stairs))

saveRDS(Stairs, file="stairs_df.rds")

```


## Combine all Data Sets

This next part combines all of the data sets including weather and the glucose levels into one data set. I also filtered out any na for the glucose level simply because this is what we are interested in studying specifically.

```{r, warning=FALSE, message=FALSE}
#Join all data
Diabetes <- left_join(Dex_Final, Heart_Rate, by='date')
Diabetes <- left_join(Diabetes, Active_Energy, by='date')
Diabetes <- left_join(Diabetes, Basal_Energy, by='date')
Diabetes <- left_join(Diabetes, Distance, by='date')
Diabetes <- left_join(Diabetes, Steps, by='date')
Diabetes <- left_join(Diabetes, Stand_Time, by='date')
Diabetes <- left_join(Diabetes, Stairs, by='date')
Diabetes <- left_join(Diabetes, Exercise_Time, by='date')
Diabetes <- left_join(Diabetes, Walking_Speed, by='date')
Diabetes <- left_join(Diabetes, Weather, by=c("date"="timestamp"))

#Remove NA glucose levels
Diabetes1 <- Diabetes %>%
  filter(!is.na(Mean_Glucose))%>%
  mutate_if(is.numeric, round, digits=2)

Diabetes1$Day <- as.Date(Diabetes1$date)
Diabetes1$Time <- format(Diabetes1$date, "%H")

#view complete dataset
datatable(Diabetes1)

```

## Find Relationships

Because all of my variables are numeric, I want to start by graphing every variable on a scatter plot against the glucose level. There doesn't appear to be anything significant from the plots. 

```{r, warning=FALSE, message=FALSE}

#There were too many to actually be able to see anything significant
#pairs(Diabetes1, col = 'darkblue', main = "Pairs Plot for All Variables")

#help with plot visualization
par(mfrow = c(3, 4))

ggplot(Diabetes1, aes(x=Mean_Heart_Rate, y=mean(Mean_Glucose)))+
         geom_point(color='darkblue', pch=18)+
        theme_bw()

ggplot(Diabetes1, aes(x=Mean_Active_Energy, y=mean(Mean_Glucose)))+
         geom_point(color='darkblue', pch=18)+
        theme_bw()

ggplot(Diabetes1, aes(x=Mean_Basal_Energy, y=mean(Mean_Glucose)))+
         geom_point(color='darkblue', pch=18)+
        theme_bw()

#This plot did not provide anything helpful, it was a vertical line at the 0
#ggplot(Diabetes1, aes(x=Distance, y=Mean_Glucose))+
        #geom_point(color='darkblue', pch=18)+
        #theme_bw()

ggplot(Diabetes1, aes(x=Steps, y=mean(Mean_Glucose)))+
         geom_point(color='darkblue', pch=18)+
        theme_bw()

ggplot(Diabetes1, aes(x=Stand_Time, y=mean(Mean_Glucose)))+
         geom_point(color='darkblue', pch=18)+
        theme_bw()

ggplot(Diabetes1, aes(x=Stairs, y=mean(Mean_Glucose)))+
         geom_point(color='darkblue', pch=18)+
        theme_bw()

ggplot(Diabetes1, aes(x=Exercise_Time, y=mean(Mean_Glucose)))+
         geom_point(color='darkblue', pch=18)+
        theme_bw()

ggplot(Diabetes1, aes(x=Mean_Walking_Speed, y=mean(Mean_Glucose)))+
         geom_point(color='darkblue', pch=18)+
        theme_bw()

ggplot(Diabetes1, aes(x=Temperature, y=mean(Mean_Glucose)))+
         geom_point(color='darkblue', pch=18)+
        theme_bw()

ggplot(Diabetes1, aes(x=Precipitation_Total, y=mean(Mean_Glucose)))+
         geom_point(color='darkblue', pch=18)+
        theme_bw()

ggplot(Diabetes1, aes(x=Humidity, y=mean(Mean_Glucose)))+
         geom_point(color='darkblue', pch=18)+
        theme_bw()

ggplot(Diabetes1, aes(x=Cloud_Cover, y=mean(Mean_Glucose)))+
         geom_point(color='darkblue', pch=18)+
        theme_bw()

ggplot(Diabetes1, aes(x=Sunshine_Duration, y=mean(Mean_Glucose)))+
         geom_point(color='darkblue', pch=18)+
        theme_bw()
```

```{r}
ggplot(Diabetes1, aes(x=Time, y=Mean_Glucose))+
  geom_boxplot(fill="dodgerblue3", color='black')+
  labs(title = "Average Glucose Level Throughout the Day", y="Glucose Level", x="Hour of the Day")+
  theme_bw()
```


```{r}
ggplot(Diabetes1, aes(x = Day, y = Mean_Glucose)) + 
  geom_point(fun = "mean", stat="summary", color="dodgerblue3", pch=18)+
  labs(title = "Average Glucose Throughout the Year", y="Glucose Level", x="Date")+
  theme_bw()
```

```{r, warning=FALSE}
ggplot(Diabetes1, aes(x = Mean_Heart_Rate, y = Mean_Glucose)) + 
  geom_point(fun = "mean", stat="summary", color="dodgerblue3", pch=18)+
  labs(title = "Average Glucose by Heart Rate", y="Glucose Level", x="Heart Rate")+
  theme_bw()
```

```{r, warning=FALSE}
ggplot(Diabetes1, aes(x = Mean_Active_Energy, y = Mean_Glucose)) + 
  geom_bar(fun = "mean", stat="summary", color="dodgerblue3")+
  labs(title = "Average Glucose by Active Energy", y="Glucose Level", x="Active Energy")+
  theme_bw()
```

```{r, warning=FALSE}
ggplot(Diabetes1, aes(x = Mean_Basal_Energy, y = Mean_Glucose)) + 
  geom_point(fun = "mean", stat="summary", color="dodgerblue3", pch=18)+
  labs(title = "Average Glucose by Basal Energy", y="Glucose Level", x="Basal Energy")+
  theme_bw()
```

```{r, warning=FALSE}
ggplot(Diabetes1, aes(x = Steps, y = Mean_Glucose)) + 
  geom_point(fun = "mean", stat="summary", color="dodgerblue3", pch=18)+
  labs(title = "Average Glucose by Steps", y="Glucose Level", x="Steps in a Day")+
  theme_bw()
```

```{r, warning=FALSE}
ggplot(Diabetes1, aes(x = Stand_Time, y = Mean_Glucose)) + 
  geom_bar(fun = "mean", stat="summary", fill="dodgerblue3")+
  labs(title = "Average Glucose by Stand Time", y="Glucose Level", x="Stand Time")+
  theme_bw()
```

```{r, warning=FALSE}
ggplot(Diabetes1, aes(x = Stairs, y = Mean_Glucose)) + 
  geom_bar(fun = "mean", stat="summary", fill="dodgerblue3")+
  labs(title = "Average Glucose by Stairs", y="Glucose Level", x="Stairs")+
  theme_bw()
```

```{r, warning=FALSE}
ggplot(Diabetes1, aes(x = Exercise_Time, y = Mean_Glucose)) + 
  geom_bar(fun = "mean", stat="summary", fill="dodgerblue3")+
  labs(title = "Average Glucose by Exercise Time", y="Glucose Level", x="Exercise Time")+
  theme_bw()
```

```{r, warning=FALSE}
ggplot(Diabetes1, aes(x = Mean_Walking_Speed, y = Mean_Glucose)) + 
  geom_point(fun = "mean", stat="summary", color="dodgerblue3", pch=18)+
  labs(title = "Average Glucose by Walking Speed", y="Glucose Level", x="Walking Speed")+
  theme_bw()
```

```{r, warning=FALSE}
ggplot(Diabetes1, aes(x = Temperature, y = Mean_Glucose)) + 
  geom_point(fun = "mean", stat="summary", color="dodgerblue3", pch=18)+
  labs(title = "Average Glucose by Temperature", y="Glucose Level", x="Temperature")+
  theme_bw()
```

```{r, warning=FALSE}
ggplot(Diabetes1, aes(x = Precipitation_Total, y = Mean_Glucose)) + 
  geom_point(fun = "mean", stat="summary", color="dodgerblue3", pch=18)+
  labs(title = "Average Glucose by Precipitation", y="Glucose Level", x="Precipitation")+
  theme_bw()
```

```{r, warning=FALSE}
ggplot(Diabetes1, aes(x = Humidity, y = Mean_Glucose)) + 
  geom_point(fun = "mean", stat="summary", color="dodgerblue3", pch=18)+
  labs(title = "Average Glucose by Humidity", y="Glucose Level", x="Humidity")+
  theme_bw()
```

```{r, warning=FALSE}
ggplot(Diabetes1, aes(x = Cloud_Cover, y = Mean_Glucose)) + 
  geom_point(fun = "mean", stat="summary", color="dodgerblue3", pch=18)+
  labs(title = "Average Glucose by Cloud Coverage", y="Glucose Level", x="Cloud Coverage")+
  theme_bw()
```

```{r, warning=FALSE}
ggplot(Diabetes1, aes(x = Sunshine_Duration, y = Mean_Glucose)) + 
  geom_point(fun = "mean", stat="summary", color="dodgerblue3", pch=18)+
  labs(title = "Average Glucose by Sunshine Duration", y="Glucose Level", x="Sunshine Duration")+
  theme_bw()
```




